---
title: "Use mixed effect model to explore the relation between drilling and Beta radiation"
author: "Longxiang Li"
date: "`r format(Sys.Date())`"
output: 
     html_document:
      toc: true
      fig_caption: yes
---

# Introduction

The primary goal of this study is to explore the relationship between background $\beta$ radiation level and the drilling activity.
Much of the petroleum and natural gas developed in the U.S. was created in the earth's crust at the site of ancient seas by the decay of sea life.
As a result, these shale, petroleum and gas deposits often occur in aquifers containing brine (salt water).
Radionuclides, along with other minerals that are dissolved in the brine, separate and settle out, forming various wastes at the surface:

* Mineral scales inside pipes.

* Sludges/sediments.

* Contaminated equipment or components.

* Produced waters.

Most of the produced waster is re-injected to the formation. But the disposal of the potential harzardous solid waste is almost unmonitored. It's 
reported in the news that the these wastes are dumped in open-air pit and residents nearby are complaining. Even though the citizens were most concerned
about the chemicals left in the waste such as benzne and the unknown risk in the fracking liquid, the radiation may also be emitted. So, the on-site
There's no statistic about the production/disposal of these solid waste. But it seems the solid waster from drilling goes to local waste management facility.
Besides the waste, the drilling wells themselves can accelerate the leaking of underground radon.

If this's the case, we may detect the variation of $\beta$ and relates it with the drilling activities.

# Data Collection and Management
In the past weeks, We have loaded all the data within our reach to the database and extract the data we need in our desired format.
Currently, we organize our information based on the RadNet monitors per month. The original measurement frequency of $\beta$ radiation is 3 days. But to
fit the data reporting frequency of drilling, we aggregated the $\beta$ radiation measurement (monthly average). Then we got the monthly $\beta$ data merged 
with gas/oil production data. This binding was based on the buffer around tha RadNet monitor. All gas/oil drilling information within this buffer was
aggregated (number of wells, sum production). Besides the general aggregation, we also did categorical aggregation based on drilling type (horizontal and vertical drilling). To 
find the optimal buffer radius, we created four cases: 25 km, 50 km, 75 km and 100 km.

These organized data files are stored in the shared dropbox folder now. In addition to the monthly $\beta$, categorized/overall production data, we
also add $PM_{2.5}$ mass data and $PM_{2.5}$ speciation data, field basin polygon data, USGS aeroradiometric data (including Potassium, Thorium and Uranium).These files
need strict check before being shared with other collaborators. Thee are some other datasets not included in the final .csv files such as the $\gamma$
measurement. 


```{r load-data, message = F, echo = F}
library(tidyverse)
library(mgcv)
library(lme4)
library(lmerTest)
library(ggplot2)
library(ggmap)
library(rpostgis)
library(RPostgreSQL)
library(raster)
library(gridExtra)
library(kableExtra)
library(formatR)
library(grid)
library(cowplot)
```

# Exploratory Analysis
We can check the data quality by visualizing it. We can have a general idea of the data collection process.

```{r connected to Fracking databse and extract location of wells,echo=F,message=F,cache=T}
pw <- {
  "koutrakis"
}
drv <- dbDriver("PostgreSQL")
con <- dbConnect(drv, dbname = "Fracking_Data",
                 host = "localhost", port = 5432,
                 user = "postgres", password = pw)
rm(pw)
us_basin<-pgGetGeom(con,"us_basin","basin_geom")
us_gas<-pgGetGeom(con,"Gas_Well_Headers","gas_well_geom")
us_oil<-pgGetGeom(con,"Oil_Well_Headers","oil_well_geom")
radnet<-pgGetGeom(con,"RadNet_Sp","radnet_geom")
```

First, we extract the location of wells and map it over basin boundaries. In the following plot, all active wells during the study period 01/01/2002-
12/21/2017 are ploted. 
We can see that the majority of gas drilling (blue points) and oil 
drilling (green points) are located within the basin (grey polygon). It's believed that the concentration of radionuclides varies remarkably
between geological formations. Adding basin as a categorical factor may help in the following study. 

```{r Visualize the location of all wells,echo=F,warning=F,message=F,cache=T,out.width='95%'}
rm(center)
center<-ggmap::geocode("United States",source = "dsk")
while(is.na(center[1])){try(center<-geocode("United States"))}
while(!exists("US_map")){try(US_map<-get_googlemap(as.numeric(center),zoom = 9,key="AIzaSyBKUgVXmsfj_HQ77IxRjSSNs7eQxYtyNis"))}
map<-ggmap(US_map)
map<-map+geom_polygon(data = fortify(us_basin),
                      aes(long, lat, group = group),
                      colour = "black", fill="grey",alpha = 0.5,size=1)
part_gas<-cbind.data.frame("Gas_Well",coordinates(us_gas))
names(part_gas)<-c("Type","lon","lat")
part_oil<-cbind.data.frame("Oil_Well",coordinates(us_oil))
names(part_oil)<-c("Type","lon","lat")
part_rad<-cbind.data.frame("RadNet_Monitor",coordinates(radnet))
names(part_rad)<-c("Type","lon","lat")
vis_data<-rbind.data.frame(part_gas,part_oil,part_rad)
map<-map+geom_point(data=vis_data,aes(x=lon,y=lat,color=Type,alpha=Type,size=Type))+
  scale_color_manual(values = c("Gas_Well" = "navyblue","Oil_Well" = "darkgreen",'RadNet_Monitor' = "Red"))+
  scale_alpha_manual(values= c("Gas_Well" = 0.05,"Oil_Well" = 0.05,'RadNet_Monitor' = 0.8))+
  scale_size_manual(values=c("Gas_Well" = 0.25,"Oil_Well" = 0.25,'RadNet_Monitor' = 1.5))
map<-map+ggtitle("Location of Gas/Oil Drillings in the Lower 48 States")
print(map)
```

Then we plot the locations of vertical drillings and horizontal drillings. Horizontal drillings are prevailing now due to its higher productivity.
Currently, the majority of new constructed drills are horizontal.But this trend doesn't occurs in every field. We can see that there're only few 
hrizontal drillingin Lousaina where the dominant drilling type is still vertical.

```{r Visualize the location of horizontal and vertical drills,echo=F,warning=F,message=F,cache=T,out.width='95%'}
map<-ggmap(US_map)
map<-map+geom_polygon(data = fortify(us_basin),
                      aes(long, lat, group = group),
                      colour = "black", fill="grey",alpha = 0.5,size=0.5)
part_gas<-cbind.data.frame("H_Gas_Well",coordinates(us_gas[us_gas$`Drill Type`=="H",]))
names(part_gas)<-c("Type","lon","lat")
part_oil<-cbind.data.frame("H_Oil_Well",coordinates(us_oil[us_oil$`Drill Type`=="H",]))
names(part_oil)<-c("Type","lon","lat")
part_rad<-cbind.data.frame("RadNet_Monitor",coordinates(radnet))
names(part_rad)<-c("Type","lon","lat")
vis_data<-rbind.data.frame(part_gas,part_oil,part_rad)
h_map<-map+geom_point(data=vis_data,aes(x=lon,y=lat,color=Type,alpha=Type,size=Type))+
  scale_color_manual(values = c("H_Gas_Well" = "navyblue","H_Oil_Well" = "darkgreen",'RadNet_Monitor' = "Red"))+
  scale_alpha_manual(values= c("H_Gas_Well" = 0.05,"H_Oil_Well" = 0.05,'RadNet_Monitor' = 0.8))+
  scale_size_manual(values=c("H_Gas_Well" = 0.25,"H_Oil_Well" = 0.25,'RadNet_Monitor' = 1.5))+
  theme(legend.position="bottom")
h_map<-h_map+ggtitle("Location of Fracking Drillings in the Lower 48 States")
h_map<-h_map+theme(plot.title = element_text(size = 6, face = "bold"))

map<-ggmap(US_map)
map<-map+geom_polygon(data = fortify(us_basin),
                      aes(long, lat, group = group),
                      colour = "black", fill="grey",alpha = 0.5,size=0.5)
part_gas<-cbind.data.frame("V_Gas_Well",coordinates(us_gas[us_gas$`Drill Type`=="V",]))
names(part_gas)<-c("Type","lon","lat")
part_oil<-cbind.data.frame("V_Oil_Well",coordinates(us_oil[us_oil$`Drill Type`=="V",]))
names(part_oil)<-c("Type","lon","lat")
part_rad<-cbind.data.frame("RadNet_Monitor",coordinates(radnet))
names(part_rad)<-c("Type","lon","lat")
vis_data<-rbind.data.frame(part_gas,part_oil,part_rad)
v_map<-map+geom_point(data=vis_data,aes(x=lon,y=lat,color=Type,alpha=Type,size=Type))+
  scale_color_manual(values = c("V_Gas_Well" = "navyblue","V_Oil_Well" = "darkgreen",'RadNet_Monitor' = "Red"))+
  scale_alpha_manual(values= c("V_Gas_Well" = 0.05,"V_Oil_Well" = 0.05,'RadNet_Monitor' = 0.8))+
  scale_size_manual(values=c("V_Gas_Well" = 0.25,"V_Oil_Well" = 0.25,'RadNet_Monitor' = 1.5))+
  theme(legend.position="bottom")
v_map<-v_map+ggtitle("Location of Conventional Drillings in the Lower 48 States")
v_map<-v_map+theme(plot.title = element_text(size = 6, face = "bold"))
grid.arrange(h_map,v_map,nrow=1)
```

# Dataset of all measurement within the basin area
We select the data within the basin area after 2012.In the following figure, we plot the trend of $\beta$ radiation in each city.
We can see it seems that, the trend of $\beta$ for each city is different. Since this's not our primary research interest, we 
can regard this as a cross random effect because the time doesn't work equally for every city.

It's also worth noting that, the duration of measruement for each RadNet monitor varies.So we need to filter the data based on the
duration. In the figure we can see that: KALISPELL,MT, MOBILE,AL, PAINESVILLE,OH and TOPEKA,KS can be excluded. 


```{r summarize the data,echo=F, message=F,warning=F}
load(here::here("data","beta_gas_oil_50.RData"))
rad_qs_zones<-rad_qs[!is.na(rad_qs$basin_name),]
rad_qs_zones<-rad_qs_zones[rad_qs_zones$YEAR>2008,]
rad_qs_zones<-rad_qs_zones[!is.na(rad_qs_zones$mass),]
rad_qs_zones<-rad_qs_zones[!rad_qs_zones$city_state%in%c("KALISPELL,MT", "CORPUS CHRISTI,TX","MOBILE,AL", "PAINESVILLE,OH","TOPEKA,KS","LUBBOCK,TX","LANSING,MI","HARLINGEN,TX"),]

summary_result<-rad_qs_zones%>%
  group_by(city_state)%>%
  summarise(min_gas_num=min(Gas_Num,na.rm = T),
            max_gas_num=max(Gas_Num,na.rm = T),
            min_gas_prod=min(Gas_Prod,na.rm=T),
            max_gas_prod=max(Gas_Prod,na.rm=T),
            min_gas_h_prod=min(H_Gas_Prod,na.rm=T),
            max_gas_h_prod=max(H_Gas_Prod,na.rm=T),
            min_gas_v_prod=min(V_Gas_Prod,na.rm=T),
            max_gas_v_prod=max(V_Gas_Prod,na.rm=T),
            min_gas_h_num=min(H_Gas_Num,na.rm = T),
            max_gas_h_num=max(H_Gas_Num,na.rm = T),
            min_gas_v_num=min(V_Gas_Num,na.rm = T),
            max_gas_v_num=max(V_Gas_Num,na.rm = T),
            min_oil_num=min(Oil_Num,na.rm = T),
            max_oil_num=max(Oil_Num,na.rm = T),
            min_oil_prod=min(Oil_Prod,na.rm=T),
            max_oil_prod=max(Oil_Prod,na.rm=T),
            min_oil_h_prod=min(H_Oil_Prod,na.rm=T),
            max_oil_h_prod=max(H_Oil_Prod,na.rm=T),
            min_oil_v_prod=min(V_Oil_Prod,na.rm=T),
            max_oil_v_prod=max(V_Oil_Prod,na.rm=T),
            min_oil_h_num=min(H_Oil_Num,na.rm = T),
            max_oil_h_num=max(H_Oil_Num,na.rm = T),
            min_oil_v_num=min(V_Oil_Num,na.rm = T),
            max_oil_v_num=max(V_Oil_Num,na.rm = T),
            min_num=min(Oil_Num,na.rm = T),
            max_num=max(Oil_Num,na.rm = T),
            mean_beta=mean(beta,na.rm=T),
            n_lead_210=length(Lead_210[!is.na(Lead_210)]),
            lead_range=paste(range(Date[!is.na(Lead_210)]),collapse = "~"),
            n_beta=length(beta),
            beta_range=paste(range(Date[!is.na(beta)]),collapse = "~"),
            gas_range=paste(range(Date[Gas_Num>0]),collapse = "~"),
            oil_range=paste(range(Date[Gas_Num>0]),collapse = "~")
  )
rad_qs_zones<-filter(rad_qs_zones,city_state%in%unique(summary_result[summary_result$max_gas_num>0|summary_result$max_oil_prod>0,]$city_state))

```{r rescale the production and number of wells data,echo=F,message=F}
rad_qs_zones$H_Gas_Prod<-rad_qs_zones$H_Gas_Prod/1000000
rad_qs_zones$V_Gas_Prod<-rad_qs_zones$V_Gas_Prod/1000000
rad_qs_zones$Gas_Prod<-rad_qs_zones$Gas_Prod/1000000

rad_qs_zones$H_Gas_Num<-rad_qs_zones$H_Gas_Num/100
rad_qs_zones$V_Gas_Num<-rad_qs_zones$V_Gas_Num/100
rad_qs_zones$Gas_Num<-rad_qs_zones$Gas_Num/100

rad_qs_zones$H_Oil_Prod<-rad_qs_zones$H_Oil_Prod/1000000
rad_qs_zones$V_Oil_Prod<-rad_qs_zones$V_Oil_Prod/1000000
rad_qs_zones$Oil_Prod<-rad_qs_zones$Oil_Prod/1000000

rad_qs_zones$H_Oil_Num<-rad_qs_zones$H_Oil_Num/100
rad_qs_zones$V_Oil_Num<-rad_qs_zones$V_Oil_Num/100
rad_qs_zones$Oil_Num<-rad_qs_zones$Oil_Num/100
rad_qs_zones$log_beta<-log(rad_qs_zones$beta)
rad_qs_zones$btpm<-rad_qs_zones$beta/rad_qs_zones$mass
rad_qs_zones$radon<-as.numeric(as.character(rad_qs_zones$radon))
```{r Clean the data based on the duration of beta measurement, echo=F, message=F,warning=F,out.width='95%',cache=T}
ggplot(data=rad_qs_zones,aes(x=Date,y=beta))+geom_point(size=0.5)+geom_smooth(size=0.5)+facet_wrap(~city_state)+
  theme( axis.text = element_text( size = 5 ),
         axis.text.x = element_text( size = 5 ),
         axis.title = element_text( size = 5, face = "bold" ),
         legend.position="none",
         strip.text = element_text(size = 5))
ggplot(data=rad_qs_zones,aes(x=Date,y=Gas_Num))+geom_point(size=0.5)+geom_smooth(size=0.5)+facet_wrap(~city_state)+
  theme( axis.text = element_text( size = 5 ),
         axis.text.x = element_text( size = 5 ),
         axis.title = element_text( size = 5, face = "bold" ),
         legend.position="none",
         strip.text = element_text(size = 5))
ggplot(data=rad_qs_zones,aes(x=Date,y=Gas_Prod))+geom_point(size=0.5)+geom_smooth(size=0.5)+facet_wrap(~city_state)+
  theme( axis.text = element_text( size = 5 ),
         axis.text.x = element_text( size = 5 ),
         axis.title = element_text( size = 5, face = "bold" ),
         legend.position="none",
         strip.text = element_text(size = 5))
ggplot(data=rad_qs_zones,aes(x=Date,y=Oil_Num))+geom_point(size=0.5)+geom_smooth(size=0.5)+facet_wrap(~city_state)+
  theme( axis.text = element_text( size = 5 ),
         axis.text.x = element_text( size = 5 ),
         axis.title = element_text( size = 5, face = "bold" ),
         legend.position="none",
         strip.text = element_text(size = 5))
ggplot(data=rad_qs_zones,aes(x=Date,y=Oil_Prod))+geom_point(size=0.5)+geom_smooth(size=0.5)+facet_wrap(~city_state)+
  theme( axis.text = element_text( size = 5 ),
         axis.text.x = element_text( size = 5 ),
         axis.title = element_text( size = 5, face = "bold" ),
         legend.position="none",
         strip.text = element_text(size = 5))
```

If we have a closer look at the city-specific data, we can see the trend of gas/oil production and construction within a radius of the city. For 
Fort Worth, TX, the gas production of horizontal drilling was actually declining. In the lower panel, dash thick line indicates the construction while
the thin solid line indicates the spuding number After the spuding peak and construction around 2012, the new
construction was almost pended due to the low oil price, the production decreased gradually. There's a single oil well spudding/construction in 2015.
After that, the oil production declined gradually.


```{r visualize a demo city Fort Worth, TX,  echo=F, message=F,warning=F,cache=T,fig.width=16,fig.height=16}
for(i in 1:5){
  rm(center)
  rm(city_map)
  city<-unique(rad_qs_zones$city_state)[i]
  city_name<-city
  center<-ggmap::geocode(as.character(city_name),source = "dsk")
  while(is.na(center[1])){try(center<-geocode(as.character(city_name),source = "dsk"))}
  while(!exists("city_map")){try(city_map<-get_googlemap(as.numeric(center),zoom = 9,key="AIzaSyBKUgVXmsfj_HQ77IxRjSSNs7eQxYtyNis"))}
  city_map<-ggmap::ggmap(city_map)
  city_buffer<-buffer(radnet[radnet$city_state==city_name,],50000)
  city_map<-city_map+geom_polygon(data = fortify(city_buffer),
                                  aes(long, lat, group = group),
                                  colour = "red",fill="red", alpha = 0.1)
  city_map<-city_map+
    geom_point(data=vis_data,aes(x=lon,y=lat,color=Type,alpha=Type),size=1)+
    scale_color_manual(values = c("Gas_Well" = "navyblue","Oil_Well" = "darkgreen",'RadNet_Monitor' = "Red"))+
    scale_alpha_manual(values= c("Gas_Well" = 0.5,"Oil_Well" = 0.5,'RadNet_Monitor' = 1))+
    #scale_size_manual(values=c("Gas_Well" = 1,"Oil_Well" = 1,'RadNet_Monitor' = 1.5))+
    theme(legend.position = "bottom")
  beta_plot<-ggplot(data=rad_qs_zones[rad_qs_zones$city_state==city,],aes(x=Date,y=beta,group=YEAR))+geom_boxplot()
  gas_plot<-ggplot(data=rad_qs_zones[rad_qs_zones$city_state==city,],aes(x=Date,y=Gas_Prod))+geom_point()
  oil_plot<-ggplot(data=rad_qs_zones[rad_qs_zones$city_state==city,],aes(x=Date,y=Oil_Prod))+geom_point()
  oil_n_plot<-ggplot(data=rad_qs_zones[rad_qs_zones$city_state==city,],aes(x=Date,y=n_oil_cons))+geom_point()
  gas_n_plot<-ggplot(data=rad_qs_zones[rad_qs_zones$city_state==city,],aes(x=Date,y=n_gas_cons))+geom_point()
  #print(city_map)
  print(cowplot::plot_grid(city_map,beta_plot,gas_plot,oil_plot,oil_n_plot,gas_n_plot))
}

fortworth_data<-rad_qs_zones[rad_qs_zones$city_state=="FORT WORTH,TX",]
h_gas_prod<-fortworth_data[,c("Date","H_Gas_Prod")]
names(h_gas_prod)<-c("Date","Gas_Prod")
h_gas_prod$Type="H"
v_gas_prod<-fortworth_data[,c("Date","V_Gas_Prod")]
names(v_gas_prod)<-c("Date","Gas_Prod")
v_gas_prod$Type="V"
gas_prod_data<-rbind.data.frame(h_gas_prod,v_gas_prod)
gas_prod_plot<-ggplot(data=gas_prod_data,aes(x=Date,y=Gas_Prod,fill=Type))+geom_bar(stat='identity')+ggtitle("Gas Production Plot")

h_oil_prod<-fortworth_data[,c("Date","H_Oil_Prod")]
names(h_oil_prod)<-c("Date","Oil_Prod")
h_oil_prod$Type="H"
v_oil_prod<-fortworth_data[,c("Date","V_Oil_Prod")]
names(v_oil_prod)<-c("Date","Oil_Prod")
v_oil_prod$Type="V"
oil_prod_data<-rbind.data.frame(h_oil_prod,v_oil_prod)
oil_prod_plot<-ggplot(data=oil_prod_data,aes(x=Date,y=Oil_Prod,fill=Type))+geom_bar(stat='identity')+ggtitle("Oil Production Plot")

h_gas_cons<-fortworth_data[,c("Date", "n_h_gas_spud","n_h_gas_cons")]
names(h_gas_cons)<-c("Date","Gas_Spud","Gas_Cons")
h_gas_cons$Type="H"
v_gas_cons<-fortworth_data[,c("Date", "n_v_gas_spud","n_v_gas_cons")]
names(v_gas_cons)<-c("Date","Gas_Spud","Gas_Cons")
v_gas_cons$Type="V"
gas_cons_data<-rbind.data.frame(h_gas_cons,v_gas_cons)
gas_cons_plot<-ggplot(data=gas_cons_data,aes(x=Date))+geom_line(aes(y=Gas_Spud,color=Type))
gas_cons_plot<-gas_cons_plot+geom_line(aes(y=Gas_Cons,color=Type),linetype=2,size=2)+ggtitle("Gas construction plot")

h_oil_cons<-fortworth_data[,c("Date", "n_h_oil_spud","n_h_oil_cons")]
names(h_oil_cons)<-c("Date","Oil_Spud","Oil_Cons")
h_oil_cons$Type="H"
v_oil_cons<-fortworth_data[,c("Date", "n_v_oil_spud","n_v_oil_cons")]
names(v_oil_cons)<-c("Date","Oil_Spud","Oil_Cons")
v_oil_cons$Type="V"
oil_cons_data<-rbind.data.frame(h_oil_cons,v_oil_cons)
oil_cons_plot<-ggplot(data=oil_cons_data,aes(x=Date))+geom_line(aes(y=Oil_Spud,color=Type))
oil_cons_plot<-oil_cons_plot+geom_line(aes(y=Oil_Cons,color=Type),linetype=2,size=2)+ggtitle("Oil construction plot")

grid.arrange(gas_prod_plot,oil_prod_plot,gas_cons_plot,oil_cons_plot,
             top = textGrob("Gas/Oil Plots of Fort Worth, TX",gp=gpar(fontsize=10,font=3)))
```

We can also have a closer look at the production around Pittsburgh, PA.

```{r visualize a demo city Pittsburgh, PA,  echo=F, message=F,warning=F,cache=T,out.width='95%'}

pittsburgh_data<-rad_qs_zones[rad_qs_zones$city_state=="PITTSBURGH,PA",]
h_gas_prod<-pittsburgh_data[,c("Date","H_Gas_Prod")]
names(h_gas_prod)<-c("Date","Gas_Prod")
h_gas_prod$Type="H"
v_gas_prod<-pittsburgh_data[,c("Date","V_Gas_Prod")]
names(v_gas_prod)<-c("Date","Gas_Prod")
v_gas_prod$Type="V"
gas_prod_data<-rbind.data.frame(h_gas_prod,v_gas_prod)
gas_prod_plot<-ggplot(data=gas_prod_data,aes(x=Date,y=Gas_Prod,fill=Type))+geom_bar(stat='identity')+ggtitle("Gas Production Plot")

h_oil_prod<-pittsburgh_data[,c("Date","H_Oil_Prod")]
names(h_oil_prod)<-c("Date","Oil_Prod")
h_oil_prod$Type="H"
v_oil_prod<-pittsburgh_data[,c("Date","V_Oil_Prod")]
names(v_oil_prod)<-c("Date","Oil_Prod")
v_oil_prod$Type="V"
oil_prod_data<-rbind.data.frame(h_oil_prod,v_oil_prod)
oil_prod_plot<-ggplot(data=oil_prod_data,aes(x=Date,y=Oil_Prod,fill=Type))+geom_bar(stat='identity')+ggtitle("Oil Production Plot")

h_gas_cons<-pittsburgh_data[,c("Date", "n_h_gas_spud","n_h_gas_cons")]
names(h_gas_cons)<-c("Date","Gas_Spud","Gas_Cons")
h_gas_cons$Type="H"
v_gas_cons<-pittsburgh_data[,c("Date", "n_v_gas_spud","n_v_gas_cons")]
names(v_gas_cons)<-c("Date","Gas_Spud","Gas_Cons")
v_gas_cons$Type="V"
gas_cons_data<-rbind.data.frame(h_gas_cons,v_gas_cons)
gas_cons_plot<-ggplot(data=gas_cons_data,aes(x=Date))+geom_line(aes(y=Gas_Spud,color=Type))
gas_cons_plot<-gas_cons_plot+geom_line(aes(y=Gas_Cons,color=Type),linetype=2,size=2)+ggtitle("Gas construction plot")

h_oil_cons<-pittsburgh_data[,c("Date", "n_h_oil_spud","n_h_oil_cons")]
names(h_oil_cons)<-c("Date","Oil_Spud","Oil_Cons")
h_oil_cons$Type="H"
v_oil_cons<-pittsburgh_data[,c("Date", "n_v_oil_spud","n_v_oil_cons")]
names(v_oil_cons)<-c("Date","Oil_Spud","Oil_Cons")
v_oil_cons$Type="V"
oil_cons_data<-rbind.data.frame(h_oil_cons,v_oil_cons)
oil_cons_plot<-ggplot(data=oil_cons_data,aes(x=Date))+geom_line(aes(y=Oil_Spud,color=Type))
oil_cons_plot<-oil_cons_plot+geom_line(aes(y=Oil_Cons,color=Type),linetype=2,size=2)+ggtitle("Oil construction plot")

grid.arrange(gas_prod_plot,oil_prod_plot,gas_cons_plot,oil_cons_plot,
             top = textGrob("Gas/Oil Plots of Pittsburgh, PA",gp=gpar(fontsize=10,font=3)))
```

Based on the visualization, we can make a basic model based on the organization of the data. 
The only fixed effects of basic model are the categorical background radon value and $PM_{2.5}$ mass.
The $m_month$ term means the number of month between the month of measurement and 01/01/2001. The cross random term assign random intercept for each basin/month pair.All the following models
will have the same random effect structure with this basic model. By adding a term concerning the natural gas/oil production, we hope that the new
model may be remarkably "better"than the basic model.

```{r Baisc model formula}
rad_gas_zones<-filter(rad_qs_zones,city_state%in%unique(summary_result[summary_result$max_gas_num>50,]$city_state))
lm_basic<-lmer(log_beta ~ radon+mass+(1|MONTH/city_state)+(1|YEAR), rad_gas_zones,REML = F)
lm_basic
```

According to the summary of this model, we can see that radon  Zone 1 has the highest radon level and Zone 3 has the lowest. 
can see the mean $\log_beta$ level at Zone 3 is lower than Zone 1. This's in agrement with our knowledge.In addition, the coefficient of $PM_{2.5}$,
is positive, in agreement with knowledge.By comparing the basic model with another "more basic" model without the main effect of $PM_{2.5}$, we can 
see the AIC of our basic model is lower than the "more basic" one. And the LRTest also suggest that the $PM_{2.5}$ mass is significant.

```{r check the main effect of pm2.5 mass}
lm_basic_0<-lmer(log_beta ~ radon+(1|MONTH/city_state)+(1|YEAR), rad_gas_zones,REML = F)
anova(lm_basic,lm_basic_0)
```

We can improve the performance of basic model by adding more secondary information. This part use the USGS radiotion data. As the LRTest shows,
These three variables improve the basic model, the AIC is smaller than the original basic model. So this'll be our new basic model.

```{r Improve the basic model by adding USGS radiation data}
lm_basic_1<-lmer(log_beta~ radon+mass+Umeans+(1|MONTH/city_state)+(1|YEAR),rad_gas_zones,REML=F)
anova(lm_basic_1,lm_basic)
lm_basic<-lm_basic_1
```

In a similar way we adding a random effect based on the basin (field name). This may be related with the main effect of USGS radiation data. 
I should ask an expert about it. But according to the LRT, adding this random effect improve the basic model. So this's the basic model now.

```{r Improve the basic model by adding basin factor data}
lm_basic_2<-lmer(log_beta~ radon+mass+Umeans+(1|MONTH/city_state)+(1|YEAR)+(1|basin_name),rad_gas_zones,REML=F)
anova(lm_basic_2,lm_basic)

lm_basic<-lm_basic_2
```


(1) we check the main effect of number of gas wells in the buffer. There linear mixed effect models are made here for total number of 
gas wells, number of active vertical gas wells and number of active horizontal gas wells. These models are compared with the basic model one by one.

```{r check the main effect of number of gas wells,out.width='95%'}
lm_gas_num <- lmer(log_beta ~ radon+mass+Umeans+Gas_Num+(1|MONTH/city_state)+(1|YEAR)+(1|basin_name), rad_gas_zones,REML = F)
anova(lm_gas_num,lm_basic,test="Chisq")
fixef(lm_gas_num)

lm_gas_v_num <- lmer(log_beta ~ radon+mass+Umeans+V_Gas_Num+(1|MONTH/city_state)+(1|YEAR)+(1|basin_name), rad_gas_zones,REML = 0)
anova(lm_gas_v_num,lm_basic)
fixef(lm_gas_v_num)

lm_gas_h_num <- lmer(log_beta ~ radon+mass+Umeans+H_Gas_Num+(1|MONTH/city_state)+(1|YEAR)+(1|basin_name), rad_gas_zones,REML = 0)
anova(lm_gas_h_num,lm_basic)
fixef(lm_gas_h_num)

lm_gas_a_num<-lmer(log_beta ~ radon+mass+Umeans+n_gas_wells+(1|MONTH/city_state)+(1|YEAR)+(1|basin_name), rad_gas_zones,REML = 0)
anova(lm_gas_a_num,lm_basic)
fixef(lm_gas_a_num)

lm_gas_h_num<-lmer(log_beta ~ radon+mass+Umeans+n_h_gas_wells+(1|MONTH/city_state)+(1|YEAR)+(1|basin_name), rad_gas_zones,REML = 0)
anova(lm_gas_h_num,lm_basic)
fixef(lm_gas_h_num)

lm_gas_v_num<-lmer(log_beta ~ radon+mass+Umeans+n_v_gas_wells+(1|MONTH/city_state)+(1|YEAR)+(1|basin_name), rad_gas_zones,REML = 0)
anova(lm_gas_v_num,lm_basic)
fixef(lm_gas_v_num)

lm_gas_both_num<-lmer(log_beta~radon+mass+Umeans+n_v_gas_wells+n_h_gas_wells+(1|MONTH/city_state)+(1|YEAR)+(1|basin_name), rad_gas_zones,REML = 0)
anova(lm_gas_both_num,lm_basic)
fixef(lm_gas_both_num)

lm_gas_cons_num<-lmer(log_beta ~ radon+mass+Umeans+n_gas_cons+(1|MONTH/city_state)+(1|YEAR)+(1|basin_name), rad_gas_zones,REML = 0)
anova(lm_gas_cons_num,lm_basic)
fixef(lm_gas_cons_num)
 
lm_gas_h_cons_num<-lmer(log_beta ~ radon+mass+Umeans+n_h_gas_cons+(1|MONTH/city_state)+(1|YEAR)+(1|basin_name), rad_gas_zones,REML = 0)
anova(lm_gas_h_cons_num,lm_basic)
fixef(lm_gas_h_cons_num)

lm_gas_v_cons_num<-lmer(log_beta ~ radon+mass+Umeans+n_v_gas_cons+(1|MONTH/city_state)+(1|YEAR)+(1|basin_name), rad_gas_zones,REML = 0)
anova(lm_gas_v_cons_num,lm_basic)
fixef(lm_gas_v_cons_num)

lm_gas_spud_num<-lmer(log_beta ~ radon+mass+Umeans+n_gas_spud+(1|MONTH/city_state)+(1|YEAR)+(1|basin_name), rad_gas_zones,REML = 0)
anova(lm_gas_spud_num,lm_basic)
fixef(lm_gas_spud_num)

lm_gas_h_spud_num<-lmer(log_beta ~ radon+mass+Umeans+n_h_gas_spud+(1|MONTH/city_state)+(1|YEAR)+(1|basin_name), rad_gas_zones,REML = 0)
anova(lm_gas_h_spud_num,lm_basic)
fixef(lm_gas_h_spud_num)

```{r check the main effect of gas production,out.width='95%'}
lm_gas_prod <- lmer(log_beta~ radon+mass+Umeans+Gas_Prod+(1|MONTH/city_state)+(1|YEAR)+(1|basin_name), rad_gas_zones,REML = F)
anova(lm_gas_prod,lm_basic,test="Chisq")
fixef(lm_gas_prod)

lm_gas_v_prod <- lmer(log_beta ~ radon+mass+Umeans+V_Gas_Prod+(1|MONTH/city_state)+(1|YEAR)+(1|basin_name), rad_gas_zones,REML = 0)
anova(lm_gas_v_prod,lm_basic)
fixef(lm_gas_v_prod)

lm_gas_h_prod <- lmer(log_beta ~ radon+mass+Umeans+H_Gas_Prod+(1|MONTH/city_state)+(1|YEAR)+(1|basin_name), rad_gas_zones,REML = 0)
anova(lm_gas_h_prod,lm_basic)
fixef(lm_gas_h_prod)
```

Oil Num

```{r check the main effect of number of oil wells}
lm_oil_num <- lmer(log_beta ~ radon+mass+Umeans+Oil_Num+(1|MONTH/city_state)+(1|YEAR)+(1|basin_name), rad_gas_zones,REML = F)
anova(lm_oil_num,lm_basic,test="Chisq")
fixef(lm_oil_num)

lm_oil_v_num <- lmer(log_beta ~ radon+mass+Umeans+V_Oil_Num+(1|MONTH/city_state)+(1|YEAR)+(1|basin_name), rad_gas_zones,REML = 0)
anova(lm_oil_v_num,lm_basic)
fixef(lm_oil_v_num)

lm_oil_h_num <- lmer(log_beta ~ radon+mass+Umeans+H_Oil_Num+(1|MONTH/city_state)+(1|YEAR)+(1|basin_name), rad_gas_zones,REML = 0)
anova(lm_oil_h_num,lm_basic)
fixef(lm_oil_h_num)

lm_oil_a_num<-lmer(log_beta ~ radon+mass+Umeans+n_oil_wells+(1|MONTH/city_state)+(1|YEAR)+(1|basin_name), rad_gas_zones,REML = 0)
anova(lm_oil_a_num,lm_basic)
fixef(lm_oil_a_num)

lm_oil_h_num<-lmer(log_beta ~ radon+mass+Umeans+n_h_oil_wells+(1|MONTH/city_state)+(1|YEAR)+(1|basin_name), rad_gas_zones,REML = 0)
anova(lm_oil_h_num,lm_basic)
fixef(lm_oil_h_num)

lm_oil_v_num<-lmer(log_beta ~ radon+mass+Umeans+n_v_oil_wells+(1|MONTH/city_state)+(1|YEAR)+(1|basin_name), rad_gas_zones,REML = 0)
anova(lm_oil_v_num,lm_basic)
fixef(lm_oil_v_num)

lm_oil_cons_num<-lmer(log_beta ~ radon+mass+Umeans+n_oil_cons+(1|MONTH/city_state)+(1|YEAR)+(1|basin_name), rad_gas_zones,REML = 0)
anova(lm_oil_cons_num,lm_basic)
fixef(lm_oil_cons_num)

lm_oil_h_cons_num<-lmer(log_beta ~ radon+mass+Umeans+n_h_oil_cons+(1|MONTH/city_state)+(1|YEAR)+(1|basin_name), rad_gas_zones,REML = 0)
anova(lm_oil_h_cons_num,lm_basic)
fixef(lm_oil_h_cons_num)

lm_oil_v_cons_num<-lmer(log_beta ~ radon+mass+Umeans+n_v_oil_cons+(1|MONTH/city_state)+(1|YEAR)+(1|basin_name), rad_gas_zones,REML = 0)
anova(lm_oil_v_cons_num,lm_basic)
fixef(lm_oil_v_cons_num)

lm_oil_spud_num<-lmer(log_beta ~ radon+mass+Umeans+n_oil_spud+(1|MONTH/city_state)+(1|YEAR)+(1|basin_name), rad_gas_zones,REML = 0)
anova(lm_oil_spud_num,lm_basic)
fixef(lm_oil_spud_num)

lm_oil_h_spud_num<-lmer(log_beta ~ radon+mass+Umeans+n_h_oil_spud+(1|MONTH/city_state)+(1|YEAR)+(1|basin_name), rad_gas_zones,REML = 0)
anova(lm_oil_h_spud_num,lm_basic)
fixef(lm_oil_h_spud_num)

lm_oil_v_spud_num<-lmer(log_beta ~ radon+mass+Umeans+n_v_oil_spud+(1|MONTH/city_state)+(1|YEAR)+(1|basin_name), rad_gas_zones,REML = 0)
anova(lm_oil_v_spud_num,lm_basic)
fixef(lm_oil_v_spud_num)



```{r check the main effect of oil production}
lm_oil_prod <- lmer(log_beta ~ radon+mass+Umeans+Oil_Prod+(1|MONTH/city_state)+(1|YEAR)+(1|basin_name), rad_gas_zones,REML = F)
anova(lm_oil_prod,lm_basic,test="Chisq")
fixef(lm_oil_prod)

lm_oil_v_prod <- lmer(log_beta ~ radon+mass+Umeans+V_Oil_Prod+(1|MONTH/city_state)+(1|YEAR)+(1|basin_name), rad_gas_zones,REML = 0)
anova(lm_oil_v_prod,lm_basic)
fixef(lm_oil_v_prod)

lm_oil_h_prod <- lmer(log_beta ~ radon+mass+Umeans+H_Oil_Prod+(1|MONTH/city_state)+(1|YEAR)+(1|basin_name), rad_gas_zones,REML = 0)
anova(lm_oil_h_prod,lm_basic)
fixef(lm_oil_h_prod)
```


---
title: "08_Mixed_Effect_Model.R"
author: "lol087"
date: "Fri Oct 05 13:42:59 2018"
---
